Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-07-21 cover_letter/response_letter.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Professional Formal Letter
% LaTeX Template
% Version 1.0 (28/12/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Brian Moses (http://www.ms.uky.edu/~math/Resources/Templates/LaTeX/)
% with extensive modifications by Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% may not be compatible with rest of template
% \documentclass[journal=jpcbfk,manuscript=article]{achemso}

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

%\documentclass[11pt,a4paper,draft]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc) 
\documentclass[11pt,a4paper]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc) 
\usepackage{letterbib}
\usepackage{graphicx} % Required for including pictures
\usepackage{microtype} % Improves typography
\usepackage{fancyhdr}
\usepackage[T1]{fontenc} % Required for accented characters
\usepackage[sort&compress,numbers,super]{natbib}
\usepackage{caption}
\usepackage{newfloat}
\usepackage{float}
\usepackage{url}
\usepackage{color}
\usepackage{soul}
\usepackage[svgnames]{xcolor}
\usepackage{framed}
\usepackage{calligra}
\usepackage{fp}
\FPset\marginWidth{0.7}	% margin width (all sides) in inches
\usepackage[margin=\marginWidth in]{geometry}
\usepackage{marginfix}
\usepackage[colorlinks=true,
            linkcolor=blue,
            urlcolor=blue,
            citecolor=blue]{hyperref}
\usepackage[font={small,sf}, singlelinecheck=false]{caption}
\usepackage{capt-of}

@@

\renewcommand{\thesfigure}{S\arabic{sfigure}}
\renewcommand{\thelfigure}{\Alph{lfigure}}

%-----------------------------------------------------------------------------
% HEADER/FOOTER INFO
%-----------------------------------------------------------------------------
%\pagestyle{fancy}
\fancyhf{}
%\cfoot{\textit{telephone} 217-333-1441 $\bullet$ \textit{fax} 217-333-2736 $\bullet$ \textit{email} matse@illinois.edu}
\cfoot{\thepage}

%----------------------------------------------------------------------------------------
%	SENDER INFORMATION
%----------------------------------------------------------------------------------------

\def\Who{Andrew L. Ferguson} % Your name
\def\What{, PhD} % Your title
\def\Where{Pritzker School of Molecular Engineering\\University of Chicago} % Your department/institution
\def\department{Pritzker School of Molecular Engineering}
\def\institution{University of Chicago}
\def\Address{5640 South Ellis Avenue} % Your address
\def\CityZip{Chicago, IL 60637} % Your city, zip code, country, etc
\def\Email{e:~\url{andrewferguson@uchicago.edu}} % Your email address
\def\TEL{p:~(773)~702-5950} % Your phone number
\def\URL{w:~\url{www.ferglab.com}} % Your URL
\def\titleone{Associate Professor\\ Vice Dean for Equity, Diversity, and Inclusion}

@@

\begin{letter}
{
Prof.\ Hanadi Sleiman \\
Associate Editor, \textit{The Journal of the American Chemical Society}
}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

\opening{Dear Prof.\ Sleiman}

Thank you for your 6-Jul-2021 message regarding our article ``Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy'' assigned manuscript ID ja-2021-05219p. We thank you for handling our submission and the anonymous reviewers for their careful reading and critical assessment of our work.

@@

We reproduce below the full text of the reviewer comments in blue along with our responses, reproduction of the changes made to our manuscript, and the locations of these edits. The modifications within our revised submission in response to the reviewer comments are indicated in red text.

In light of these changes, we hope that our revised submission is suitable for publication in \textit{The Journal of the American Chemical Society}.

@@

\end{letter}

%----------------------------------------------------------------------------------------

\clearpage
\newpage

\noter{Additional typo to fix on p.~18: ``H, 5S2, 3S2, 5S4, 3S4, F'' -> ``H, 5S2, 3S2, 5S4, 3S4, F4''. Probably no need to mention it as this is trivial.}

@@

\begin{shaded}
\textbf{Response to Reviewer \# 1}
\end{shaded}

\textcolor{blue}{In the manuscript ``Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy'', Jones et al. provide a well-designed and compelling study of DNA hybridization kinetics and dynamics that succinctly encompasses both experiment and theory. They demonstrate that the coarse-grained model of DNA, 3SPN.2, can not only well-reproduce the relative kinetic rates of sequence-dependent hybridization but also can reveal critical insights into their physical underpinnings. These insights are driven by the use of state-of-the-art applications of machine learning to Markov state model construction and well-validated with temperature jump infrared spectroscopy. In particular, the demonstration of how the placement of two G:C base pairs within the sequence can dramatically affect the energy landscape is truly impactful for the design of DNA sequences with desired properties. We believe this work is well suited for JACS publication after the Authors address a few major and minor points.}

\textbf{Author Reply:}   We thank the reviewer for their close reading of the manuscript and are pleased that they find the work impactful and suitable for publication in JACS.

\textcolor{blue}{Major point.
1)      Many of the core conclusions about the hybridization and dehybridzation mechanisms are based on the designation of the relevant metastable states as H, 5S2, 3S2, 5S4, F4, and D. While the authors do an excellent job of demonstrating that 2-6 state models provide an accurate description of the dynamics, in the current form, the paper does not provide sufficient evidence to support their chosen representative microstates and associated nomenclature (the only evidence being the ``representative'' structures shown in Figures 2A, C). A given macrostate represents an ensemble of microstates and it should be demonstrated that a significant fraction of the microstates that have a high probability of belonging to a particular macrostate share the same or similar base-pairing pattern. This would better justify how the authors chose to describe these states as 5S2, 3S2, etc., and provide the reader with more confidence that the mechanisms follow from the simulations. At the least, the authors should describe how the representative microstates (depicted in Figure 2A, C) were obtained.
}

\textbf{Author Reply:} The designation and structural representation of the macrostates is crucial to the core conclusions of the manuscript and we concur that we can and should better expose for the reader how the representative microstates were selected, how representative they are of the microstate ensemble within the macrostate, and our analysis of the base-pairing patterns within each macrostate that motivated our naming scheme.

%\noteg{I think this comment is really getting at why the cartoon representations are appropriate to describe the state as a whole and is less concerned with how we obtained the molecular rendering. I've tried to address both aspects below and can add more detail the first figure or choose a different kind of representation.}

On \rood{p.~17}, we have added the following text:

``...each of these macrostates. \rood{The representative microstates emblematic of each macrostate were selected randomly from those possessing high ($>$99\%) macrostate membership probabilities. Since the microstate ensemble exhibits a range of conformations, however, it is useful to characterize the degree of structural heterogeneity to determine the degree to which these microstates are emblematic of the distribution. In Fig.~S{\blauw{X}} we present a projection of our macrostates into two physical order parameters -- the degree of 3$^\prime$ shift and degree of 5$^\prime$ shift -- to provide an interpretable embedding of the macrostates that exposes base pairing patterns and structural heterogeneity within the microstate ensemble. In all cases, we find these microstate distributions to be relatively narrowly focussed within the low-free energy core of each macrostate such that a single archetypal microstate is indeed a good representative for the ensemble and an accurate representation of the base pairing pattern of the macrostate.} In Fig. 2b we present...''

\noter{Fig. SX should contain four panels: (a) Explication of the two order parameters. (b) Current macrostate\_colors\_FES but with improved schematics on x and y axes. Use $d_{3^\prime}^{-1}$ and $d_{5^\prime}^{-1}$ as axis labels. Please also provide a numerical scale for the free energy color map. Perhaps also draw a contour line around each free energy well showing the region that contains, say, 90\% of the probability mass -- the idea is that we want to show that the microstate ensembles are very much contained within a tight distribution at the FE minima. (c) Same as (b) but showing and heat map showing not FE but state membership probabilities. (d) Table listing mean and standard deviation for 3' shift and 5' shift for each state in each system in both (i) real units and (ii) reciprocal units, and the last column showing a superposition of $\sim$50 representative configurations with $>$90\% membership probabilities to try to show the structural variation in the ensemble.}

We also added the new Fig.~S\rood{X} to the Supporting Information where we discuss at some length within the caption the important issues raised by the reviewer. The figure and its caption are reproduced below.

\noter{When reproducing a main text or SI figure in the response letter, I think it's a good practice to typically (i) reproduce the whole figure and caption and (ii) give it the same numbering as in the manuscript rather than an independent numbering just for the response letter. Figures generated just for the response letter can have some other numbering scheme (e.g., R1, R2, ...) or left without numbers.}

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=150mm]{./revision_figures/macrostate_colors_FES.png}
        %\caption{Caption will probably need to go on next page -- big figure and caption. Might need multipage caption: https://tex.stackexchange.com/questions/399698/how-to-get-figure-caption-to-span-multiple-pages-without-having-to-switch-every}
        \label{fig:macrostate_colors_FES}
	\end{center}
\end{figure}
\clearpage
%\newpage
\begin{figure}[ht!]
	\begin{center}
        %\includegraphics[width=150mm]{./revision_figures/macrostate_colors_FES.png}
        \caption*{\textbf{Figure S\rood{X}.} Macrostate interpretation and structural analysis. The Markov state model macrostates are learned from the simulation trajectories using the featurization and deep learning procedure described in Section 2.1.2 of the main text, but it is informative to perform a \textit{post hoc} projection of the simulation data into a low-dimensional space spanned by physically interpretable order parameters in order to determine the structural characteristics of each macrostate and the structural heterogeneity of the microstate ensemble. (a) Definition of two physically-interpretable order parameters useful in characterizing the structure and base pairing of each microstate under both shifting and fraying. We define the $3^\prime$ shift for a particular base on a strand as the linear distance between that base and the base on the complementary strand offset by two base pairs in the $3^\prime$ direction. This distance reaches a minimum when the two strands are mutually shifted two base pairs in the $3^\prime$ direction. We define $d_{3^\prime}$ as the mean $3^\prime$ shift averaged over all bases in the complementary strands. It is more convenient to work the reciprocal $d_{3^\prime}^{-1}$, since it has a finite range whereas $d_{3^\prime}$ diverges for dehybridized strands. We define $d_{5^\prime}^{-1}$ analogously as the reciprocal mean $5^\prime$ shift. We have found this pair of variables extremely informative in revealing the microstate and macrostate structure and base pairing patterns. (b) Superposition of free energy surfaces for each macrostate for each sequence AT-all, GC-end, GC-core, and GC-mix projected into $d_{3^\prime}^{-1}$-$d_{5^\prime}^{-1}$. This panel conveys two important pieces of information. First. each macrostate is anchored around a single free energy minimum containing the preponderance of the probability mass of the microstate ensemble. Dashed lines indicate the contour around the free energy minimum containing 90\% of the microstate ensemble. These projections indicate a relatively narrowly peaked distribution of microstates around the free energy minimum with a broad tail. Second, the location of each macrostate free energy minimum in the $d_{3^\prime}^{-1}$-$d_{5^\prime}^{-1}$ projection readily illuminates the base pair pattern within the duplex. Within the fully hybridized H state (blue) $d_{3^\prime}^{-1}$ $\approx$ $d_{5^\prime}^{-1}$ $\approx$ 1.2 nm$^{-1}$. A 2 b.p.\ shift to 3$^\prime$ is manifest in a decrease in $d_{3^\prime}^{-1}$ (increase in $d_{3^\prime}$) and a an increase in $d_{5^\prime}^{-1}$ (increase in $d_{5^\prime}$) producing the stable 3S2 state (green). A further 2 b.p.\ shift to 3$^\prime$ results in a further decrease in $d_{3^\prime}^{-1}$ and also a decrease in $d_{5^\prime}^{-1}$ (back down to the same value as in the H state) to produce the 3S4 state (black). Analogous arguments apply to the 2 b.p.\ 5$^\prime$ shifted state 5S2 (orange) and 4 b.p.\ 5$^\prime$ shifted state 5S4 (purple). The infinitely separated dehybridized state D exists at the origin, but limited by the finite box size corresponds to $d_{3^\prime}^{-1}$ $\approx$ $d_{5^\prime}^{-1}$ $\approx$ 0.2 nm$^{-1}$ (red). The GC-core frayed state F4 (orange) exists as a heart-shaped region between the H and D sates, with each symmetric lobe of the heart corresponding to fraying of one or other of the loose ends. (c) Coloring the microstates by their membership probabilities to their assigned macrostate under a fuzzy clustering procedure shows that microstates lying in the vicinity of the free energy minima have high probabilities whereas those at the macrostate cluster boundaries have less definitive assignments. The former comprise the preponderance of the microstate ensemble while the latter represent intermediate configurations that are transiently occupied as the system transitions from one macrostate to another. (d) We report the mean and standard deviation of $d_{3^\prime}$ and $d_{5^\prime}$ taken over all microstates comprising each macrostate for each sequence. The small standard deviations that are consistent with a narrow distribution of the bulk of the microstates around the free energy minimum of each macrostate. A selection of \rood{50} microstates selected at random from those microstates with $>$90\% membership probability provide further qualitative support of a relatively homogenous structural distribution, meaning that a single high membership probability microstate is archetypal of the preponderance of microstates comprising the macrostate.}
        %\label{fig:macrostate_colors_FES}
	\end{center}
\end{figure}

%When defining macrostates, we both visually inspected configurations and compared distributions along a variety of structural features.  During the analysis process, we have found that two distance features in particular allow for interpretable visualization of each of the seven macrostate. Specifically, these features correspond to the reciprocal mean distance between 2-base-shifted base pairs in the 5$\prime$ and 3$\prime$ directions. These features highlight not only shifted states in both directions, but also a distinct frayed regions between the H and D wells. We have visualized these features below (Fig.~\ref{fig:macrostate_colors_FES}), and use the PyEMMA free energy function to plot the distribution of each macrostate/sequence on the low dimensional feature representation. Each color map corresponds to a different macrostate according to the legend, and lighter regions indicate a higher probability at equilibrium (lower free energy). We agree with the reviewer that it is important for readers to understand the physical definitions and distributions of each macrostate and its associated microstates, and we have added this figure to the supplemental information.

%In order to demonstrate that configuration in a given macrostates are well-defined by the representations in figure 2, we show the distribution of several structural metrics for each macrostate and sequence combination. These metrics are a collection of mean intermolecular distance that we have found work well to differentiate between states.

%With regard the molecular renderings shown in figures 2A and 2C, these were chosen from microstates with the highest probability of being assigned to their corresponding macrostates. Accordingly, they should correspond to structures  that are most representative of that macrostate as whole. As the reviewer points out, it is unclear how these selected microstates were defined in the main text, and we include the following addition to section 3.2 to make this point explicit:

%``We present these seven macrostates in Fig. 2a along with schematic and cartoon renderings of representative microstates contained within each of these macrostates. \rood{Molecular renderings were sampled from microstates with the highest probability of being clustered into their respective PCCA+ macrostates and should therefore reflect the macrostate population as whole. A more detailed representation of macrostate membership is shown in supplemental figure \noteg{add supplemental reference here}''}

@@

\textcolor{blue}{1)      It is unclear how five datasets were selected used for cross-validation in the SRVs (P. 10 L. 43). One would naively think a 50:50 split of a single trajectory would result in only one dataset. A bit more detail could be provided here.}

\textbf{Author Reply:}   Five-fold cross-validation was conducted by constructing five random partitions of the 40 independent trajectories into a training set of 20 trajectories and a validation set of 20 trajectories. This manner of splitting the data at the level of complete trajectories -- as opposed to splitting a single trajectory into multiple subtrajectories -- is desirable as it allows our SRV/MSM approach to learn over long 25 $\mu$s trajectories and infer slowly relaxing collective modes that may not be accessible within short sub-trajectories~\citep{Sidky2019High-ResolutionVAMPnets}. We have better clarified this procedure for the reader by revising the text on \rood{p.~10} to read:

We guarded against overfitting using five-fold cross-validation in which we \rood{constructed five random partitions of the 40 independent simulation trajectories into a training set of 20 trajectories and a validation set of 20 trajectories}.\citep{Sidky2019High-ResolutionVAMPnets}

\textcolor{blue}{2)      For the validation of embedding dimensions (P. 11 L. 3), it is hard to judge the presence of a kink at 5D in the AT-all sequence since only a maximum of 6D is shown in Fig.~S1. Could a few additional dimensions be provided?}

\textbf{Author Reply:}   We agree that it is more difficult to see the kink for the AT-all sequence than the other sequences. As suggested by the reviewer, we added two additional modes to the cross-validation procedure that does indeed better expose this kink at five slow modes. We have updated the corresponding panel Fig.~S\rood{X} and also reproduce it below.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=75mm]{./revision_figures/AT-all_added_modes.png}
        \caption{Cross-validation for the AT-all sequence with two additional modes to show the presence of a kink in the VAMP-2 score after the fifth mode.}
        \label{fig:AT-all_crossval}
	\end{center}
\end{figure}

\textcolor{blue}{3)      The values associated with the ``slow'' and ``fast'' rates appear to be inverse quantities, but this is not mentioned in the text (P. 14 L. 44).}

\textbf{Author Reply:}   The reviewer is correct that the rates have units s$^{-1}$ and describe the inverse of the experimental timescales shown on page 14. We agree that the sentence on (P. 14 L. 44) is unclear in its current form and have made the following change:

### 2021-07-21 cover_letter/response_letter.tex

``We conducted T-jump IR experiments for each DNA sequence as a function of temperature and extracted \rood{the ``slow'' $k_d^\mathrm{slow}$  and ``fast'' $k_d^\mathrm{fast}$ rates, corresponding to processes on the 10-30 $\mu$s and 70-100 ns timescales, respectively''}

\textcolor{blue}{4)      In Figure 1 the units for rates are not provided.}

\textbf{Author Reply:}   Thank you for catching this error. Fig.~1 has been updated to include units in s$^{-1}$. \noter{Probably this is all we need and no need to reproduce the figure for such a small change.}

@@

\textcolor{blue}{5)      In the comparison between experiments and theory for slow and fast T-jump IR responses (P. 17 L. 9), were there any correlations between the goodness-of-fit in the calculated rates (Figure S4) and the agreement with experiments for a given system? It is hard to judge as is because the goodness-of-fit values were not provided with Figure S4.}

\textbf{Author Reply:}     We have updated all panels of Fig.~S4 plots to include $R^2$ goodness-of-fit values determined by \rood{linear least squares fits of the models in log-space (i.e., $\log \left( f \right) = -k_d t$}) and have also re-plotted the data on log-linear axes for easier comparisons between the data and fits. The revised figure is reproduced below. In all cases we report good fits with $R^2$ $>$ \rood{0.88}. We see marginally less good fits for the GC-core sequence compared to the other three systems, possibly -- as we observe in Section 3.1 of the main text -- due to its high propensity to fray, but the reduction in the quality of the fit is only marginal and in absolute terms the fits are still very good (\rood{X.XX} $<$ $R^2$ $<$ \rood{X.XX}).

\noter{Reproduce revised Fig.~S4 here. Please report R2 values to only two sig figs -- you also have some nans that need to be corrected. Please append the following sentence to the caption in red text: ``We report within the figure legend to each panel the coefficient of determination $R^2$ for a least squares linear fit of the model to the data in log space (i.e., $\log \left( f \right) = -k_d t$) and data are plotted on log-linear axes to facilitate visual comparison of the fits. In all cases we observe excellent fits of the models to the data with all $R^2$ values equal to or better than \rood{0.XX}.''}

%\begin{figure}[ht!]
%	\begin{center}
%        \includegraphics[width=120mm]{./revision_figures/goodness_of_fit.png}
%        \caption{Exponential fits for both slow (dissociation) and fast (fraying) response for all four sequences during ``computational T-jump'' experiments. We have updated this supplemental figure to include goodness-of-fit values determined by the $R^2$ values of log fits and plotted both the slow and fast fits on a log scale.}
%        \label{fig:goodness-of-fit}
%	\end{center}
%\end{figure}

\textcolor{blue}{6)       The authors chose to compute the temperature jump relaxation rates from fits to quantities obtained directly from the MD simulation. However, it is not clear why the authors did not also compute the relaxation of observables from estimated Markov state models following (P. N. A. S. USA. 2011;108(12):4822-4827). This approach has also been previously used to compare kinetics obtained from Markov state models with other spectroscopic temperature-jump based experiments in DNA (J. Phys. Chem. B. 2019;123(10):2291-2304.) and RNA (J. Chem. Theory Comput. 2017;108(2):926-934) oligomers which should be mentioned. If possible, it would be interesting to compare or discuss the differences between their approach and these previous approaches.}

\textbf{Author Reply:}   \noter{What do you think of this section? This, and everything else, is, of course, up for discussion and revision.}The reviewer raises an important point regarding the dynamical fingerprinting methodology that have been developed to compare MSM predictions and relaxation experiments. We are familiar with these techniques, are currently applying them in our current work on abasic DNA, and did actually attempt to apply them in the present work. These analyses, however, were less informative than we had hoped due to three primary factors.

First, the experimental data contains fast responses occurring on time scales of $\sim$100 ns. As detailed in Section 3.1 and Figure 1, these fast motions occur in our simulations at an acceleration factor of approximately 120$\times$ and so proceed on a time scale of $\sim$1 ns. As such, these motions lie below the $\tau$ = 1.2 ns lag time used to construct our MSMs that is required to converge the leading slow modes and produce a Markovian model capable of passing the Chapman-Kolmogorov test (cf. Figs.~S\rood{2-3}). As such, we cannot draw comparisons of our MSM against the fast experimental data.

Second, the dynamical fingerprinting method provides an elegant means to compare experimental data against an MSM constructed at the same state point (e.g., temperature, pressure, slat conditions). We constructed only a single MSM for each sequence at the melting temperature of the 3SPN.2 model that required the collection of 1 ms of simulation data for each sequence in order to observe sufficiently many hybridization/dehybridization events for robust fitting of the model parameters (Section 2.1.1). The experimental relaxation data, however, was collected over a range of temperatures and we wished to make comparisons across this entire range. Fitting additional MSMs was too computationally burdensome due to the high volume of data required and the cost of these simulations, so instead we conducted 120 $\times$ 1 $\mu$s simulations for each DNA sequence at each temperature. These trajectory data were too sparse to fit a robust MSM, but did allow us to make a direct comparison between the MD simulations and the experimental measurements over the entire temperature range.

We therefore elected to structure our manuscript such that we first test the MD simulation predictions against experimental data in Section 3.1, then, after having validated the simulation data upon which the MSMs are constructed, perform MSM construction and analysis in Sections 3.2-6. Finally, in Section 3.7 we close with a test of a specific prediction of our MSM (the existence of long-lived metastable shifted states) that was amenable to comparison against T-jump IR.

We concur with the reviewer that we can and should explicitly mention dynamical fingerprinting approach and provide some discussion for the reader why we did not employ this approach in the present study. To this end, we have added the following text to the end of Section 3.1:

``fast kinetics of the four DNA oligomers\rood{, and lends confidence in the use of these data for the parameterization of Markov state models of the long-time dynamics of each sequence. We observe that the dynamical fingerprinting approach presents an elegant means to compare experimental relaxation data directly against a Markov state model extracted from simulation data in terms of a ``fingerprint'' of peaks with amplitudes and time scales related to the relaxation of particular system observables~\citep{Noe2011DynamicalExperiments}. These techniques have been previously employed to in applications to base stacking of DNA dinucleoside monophosphates~\citep{Remington2019MolecularKinetics} and RNA~\citep{Pinamonti2017}. We explored the use of this approach to validate the MD simulation data but encountered challenges in resolving fast dynamical motions below the lag time of the fitted MSM and the extremely high computational cost of collecting sufficient simulation data to fit MSMs at each temperature for which experimental data was collected. Accordingly, we instead elected to perform a direct comparison between the experimental data and MD trajectories to validate the simulations themselves, then proceed to train MSMs over these data and conduct analysis and experimental tests of the MSM predictions.}''

%\noteg{This paragraph may not be necessary}
%We have since experimented with some of these techniques using built-in PyEMMA functions, and although the results are interesting, they do not provide the most informative analysis of these particular dynamics or a strong experimental connection. We believe this is the case for the following reasons: i) shifting dynamics occur on timescales approaching that of dissociation and ii) shifting populations are small in amplitude relative to fraying and full dissociation and iii) a shifting signal is difficult to interpret experimentally without a more advanced technique such a as isotope labeling. \noteg{check with Brennan here} The combined result of these effects is that shifting dynamics are thought to make small contributions blended with the slow dissociation response. We support this hypothesis with data shown in Fig.~\ref{fig:responses_all}) \noteg{not sure how much to go into here}.

%\begin{figure}[ht!]
%	\begin{center}
%        \includegraphics[width=150mm]{./revision_figures/responses_all.png}
%        \caption{a) Experimental rate representations for each sequence. Both the 1600 cm$-1$ corresponding to A,T dynamics and 1660$-1$ corresponding to G,T dynamics are shown. b) Simulated relaxation using MSM built from each sequence. An A:T and G:C response are calculated based on the distance between available WC pairs. c) Change in response when factoring in shifting motions. Because shifting is low in amplitude and occurs on timescales approaching full dissociation, the response is not expected to be clearly discernible experimentally.  \noteg{need more description here, but not sure how much of this to include} }
%        \label{fig:responses_all}
%	\end{center}
%\end{figure}

% Can also describe fraying issue
% The appraoch in PNAS assumes that the dynamical observable can be captured by a markovian lag time
%The second problem we encountered with the MSM relaxation approach is that it does not adequately describe fast-termini fraying. As discussed in section 3.1, the experimental fast responses is thought to corresponds to single base fraying of (mostly A:T) termini. This is a process that occurs even faster in simulation as shown by our derived acceleration factor. We found that although fraying does occur for each sequence, this process cannot be captured by a Markovian lag time. On the other hand, multi-base fraying observed in GC-core is slower and corresponds to the second leading slow mode for that sequence. As shown in Fig.~\ref{fig:responses_all}b, we only see a distinct fast response for GC-core despite the fact that termini fraying is occurring in each sequence (as demonstrated by Fig.~1 in the main text).

%In addition to the issues discussed above, the main reason that we extracted responses directly from MD simulations was that effects of coarse graining, concentration, salt, etc. were unknown and could not be accounted for at a single temperature. For the slow response in particular,  timescales have an exponential dependence on temperature which could not be sufficiently standardized between experiment and simulation. Furthermore, generating equilibrium simulation data over a range of temperatures for each sequence was computationally prohibitive, prompting our decision to run an ensemble of shorter simulations at each temperature.

%The additional references cited by the reviewer highlight important related work and should be included \citep{Noe2011DynamicalExperiments, Pinamonti2017, Remington2019MolecularKinetics}. Although we discussed our work's connection to (J. Chem. Theory Comput. 2017;108(2):926–934)\citep{Pinamonti2017} in the introduction, we did not originally reference it with respect to relaxation rate simulations. We do believe that our work is quite distinct from these previous studies, given that much smaller RNA oligonucleotide systems were investigated. Furthermore, the RNA studies were either lacking experimental data or compared against older data which suffered some experimental flaws. This highlights the difficulty in making strong experimental comparions for oligonucleotide systems using MSM relaxation techniques alone.

%We have added the following text to section 3.1 make some of these difficulties more explicit and qualify our approach:

%\rood{"Previous studies have used a dynamical fingerprinting approach \citep{Noe2011DynamicalExperiments} to compute rates from MSMs. These techniques were used to study base stacking dynamics in 2-4 base RNA oligomers \citep{Pinamonti2017, Remington2019MolecularKinetics} and some attempts were made to compare against experiment. However, Limitations in all-atom force fields and potential experimental artefacts have made strong comparisons difficult to achieve for oligonucleotide systems \citep{Pinamonti2017}. We tried applying these techniques with our SRV-MSMs as well, but found that i) termini fraying timescales were too fast for a Markovian lag time and ii) coarse-graining effects could not be properly standardized between experiment and simulations at a single temperature. This motivated us to directly analyze responses from simulation data over a range of temperatures."}

@@

\begin{shaded}
\textbf{Response to Reviewer \# 2}
\end{shaded}

\textcolor{blue}{This paper reports simulation results on the hybridization and dehybridization mechanisms near to the melting temperature of a number of short DNA sequences, and accompanying temperature jump IR experiments. The sequences are chosen to have a repeating AT motif that potentially allows relatively stable out-of register duplex states, with a focus being on how these states are disrupted, by the presence of GC base pairs. The authors show that the introduction of just 2 appropriately-positioned base pairs can change the kinetics to a simple two-state behaviour from one in which multiple states contribute.}

\textcolor{blue}{I found this an interesting study that nicely illustrated some of the potential effects of sequence on (de)hybridization mechanisms. However, I found the results unsurprising, and these types of effects have been previously explored in the literature. Although I can't therefore recommend publication in JACS, I think it would be very appropriate for one of the more specialized ACS journals.}

\textbf{Author Reply:}  We are pleased that the reviewer finds our results interesting and suitable for publication. We concur with the reviewer that this work builds upon prior studies in the literature on sequence effects upon hybridization/dehybridization mechanisms, but goes beyond these studies in presenting a mean to rigorously resolve and quantify these dynamics. In response to comments from Reviewer \#2 and Reviewer \#3, we have updated several introductory sentences to clarify the novelty and impact of this work like less in the discovery of the particular metastable states, but in furnishing quantitative and predictive kinetic models of the dynamical transition network between these states that furnishes experimentally testable predictions and new understanding and quantification of the influence of sequence upon thermodynamics and kinetics.

We better contextualize the scope and impact of our work in the final sentence of the Abstract:

``Our results establish new understanding of the dynamical richness of sequence-dependent kinetics and mechanisms of DNA hybridization/dehybridization \rood{by furnishing quantitative and predictive kinetic models of the dynamical transition network between metastable states}, present a molecular basis with which to understand experimental temperature jump data, and furnish foundational design rules by which to rationally engineer the kinetics and pathways of DNA association and dissociation for DNA nanotechnology applications.''

and we state in the Introduction:

``Out-of-register ``shifted'' base paired structures \citep{Flamm2000RNAResolution, Romano2013DNADependence, Hinckley2014Coarse-grainedEffects, Maciejczyk2014DNAModel, Araque2016LatticeCooperativity, Xiao2019} and frayed structures \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix} stand as candidates for metastable states with the potential to mediate substantial deviations from all-or-nothing behavior, but the degree to which these states are kinetically relevant is difficult to determine experimentally and is likely to be highly sequence-dependent.''

``We validate the new computational models of hybridization/dehybridization dynamics developed in this work against new experimental data and reinterpret our prior experimental observations in light of the new computational understanding. Consistent with previous studies, \citep{Hinckley2014Coarse-grainedEffects,Romano2013DNADependence,Araque2016LatticeCooperativity} we find the degree of repetitiveness in the sequence -- and therefore the kinetic accessibility and thermodynamic stability of out-of-register shifted states -- leads to richer dynamics populated by a diversity of long-lived metastable states. Our data-driven modeling and analysis rigorously quantifies these behaviors and furnishes accurate predictive models of the hybridization/dehybridization rates, dynamical pathways, and metastable states. Specifically, we demonstrate that disrupting repetitive stretches of A:T bases by placement of interrupting G:C base pairs enables us to tune the landscape from rich six-state to simple two-state ``all-or-nothing'' behavior, and the specific location of the interrupting pair can be used to modulate the stability of long-lived frayed states.''

``Taken together, our analyses establish new molecular-level understanding of the sequence-dependent kinetics and pathways through quantitative predictive models for the long-time system dynamics, resolution of the dynamical folding pathways and metastable states, and elementary design rules with which to control the dynamical behaviors of the system. We anticipate that this new foundational understanding, and the extension of our approach to more extensive families of DNA sequences, can guide the rational design of DNA oligomers with tailored kinetic properties engineered for DNA nanotechnology applications such as DNA-PAINT. \citep{Shah2019, Strauss2020UpDNA-PAINT}''

\textcolor{blue}{One limitation of the study is its reliance on ``brute-force'' simulations (rather than using advanced sampling/rare-event techniques) to probe the transition. This means that systems of very high concentration (7 M) and temperatures close to the melting point have to be used, which are probably not the conditions of most interest.}

\textbf{Author Reply:}    (I) We agree with the reviewer that it may be possible to use biasing techniques in order to efficiently collect the data required to construct the MSMs. We are very familiar with the use of such techniques: indeed, another major interest of our research group is iterative strategies to perform data-driven determination of collective variables from simulation trajectories, enhanced sampling in these variables, and then estimation of new and improved collective variables from the enhanced sampling trajectories \noter{Add refs: 74, 66, 61, 60, 56, 44, 43 from https://www.ferglab.com/publications/}. In the present work, we chose not to employ biased simulations and instead collect unbiased trajectories for three main reasons.

-- First, the 3SPN.2 coarse-grained DNA model is sufficiently inexpensive that it was possible to collect adequate unbiased simulation data to parameterize robust MSMs at reasonable costs. As we state on \rood{p.~7}:

``Each simulation was conducted for 26 $\mu$s and frames saved to disc every 100 ps. Each simulation required $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.''

-- Second, the unbiased simulation data are in good agreement with trends in experimental T-jump IR measurements of fast and slow relaxation responses (Section 3.1, Fig.~S\rood{4}) and are sufficient to furnish robust, low-variance, and well-converged MSMs (Section 3.2, Figs.~S\rood{1-3}). This indicates that the data are of sufficient quality and volume to present favorable experimental comparisons and learn reliable long-time kinetic models.

-- Third, although it is straightforward to correct biased simulation data to recover the unbiased thermodynamics, it is more challenging to recover the unbiased kinetics. Although a number of approaches do exist, it has been our experience that the kinetic corrections can present difficulties in numerical convergence. That is not to say that we could not have successfully applied this methods, but just that we did not have cause to do so since the unbiased data were not unreasonably expensive to collect and showed favorable experimental comparisons and produced robust MSMs.

We did note the possibility of performing biased simulations in the Introduction on \rood{p.~4}, and have now elaborated these remarks to better explain our choice to proceed with unbiased simualations:
